quollr: An R Package for Visalizing 2D Models in High Dimensional Space

### 2024-02-21 paper-quollr.tex
Added shift hex grid origin to main function
Attempt 1 to have regular hexagons, still need to find out which hex_size and buffer_size need to initialize, needed to restrict the user inputs for hex_size and buffer_size because too large hex_size and buffer_size are not compatibel with data range

@@

--\textgreater{}

@@

\begin{verbatim}
num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, 
                                         x = "UMAP1", hex_size = NA)
num_bins_x
\end{verbatim}

\begin{verbatim}
#> [1] 4
\end{verbatim}

\begin{verbatim}
num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, 
                                         y = "UMAP2", hex_size = NA)
num_bins_y
\end{verbatim}

\begin{verbatim}
#> [1] 8
\end{verbatim}

@@

\begin{itemize}
\tightlist
\item
  First compute hex grid bound values along the x and y axis and generate the all the points wthin the hex box
\end{itemize}

\begin{verbatim}
cell_area <- 1

hex_size <- sqrt(2 * cell_area / sqrt(3))

buffer_size <- hex_size/2

x_bounds <- seq(min(s_curve_noise_umap[["UMAP1"]]) - buffer_size,
                  max(s_curve_noise_umap[["UMAP1"]]) + buffer_size, length.out = num_bins_x)

y_bounds <- seq(min(s_curve_noise_umap[["UMAP2"]]) - buffer_size,
                max(s_curve_noise_umap[["UMAP2"]]) + buffer_size, length.out = num_bins_y)

box_points <- expand.grid(x = x_bounds, y = y_bounds)

ggplot() +
  geom_point(data = box_points, aes(x = x, y = y), color = "red")
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf}

\begin{itemize}
\tightlist
\item
  Second for each x-value, find which y values are in the even row
\end{itemize}

\begin{verbatim}
 box_points <- box_points |>
    dplyr::arrange(x) |>
    dplyr::group_by(x) |>
    dplyr::group_modify(~ generate_even_y(.x)) |>
    tibble::as_tibble()

ggplot() +
  geom_point(data = box_points,
             aes(x = x, y = y, colour = as.factor(is_even)))
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf}

\begin{itemize}
\tightlist
\item
  Then, shift the x values of the even rows
\end{itemize}

\begin{verbatim}
## Shift for even values in x-axis
x_shift <- unique(box_points$x)[2] - unique(box_points$x)[1]

box_points$x <- box_points$x + x_shift/2 * ifelse(box_points$is_even == 1, 1, 0)

ggplot() +
  geom_point(data = box_points, aes(x = x, y = y), color = "red")
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-1.pdf}

\begin{verbatim}
all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
                                                 x = "UMAP1", y = "UMAP2", 
                                                 num_bins_x = num_bins_x, 
                                                 num_bins_y = num_bins_y, 
                                                 buffer_size = NA, hex_size = NA)

@@

\begin{verbatim}
#> Rows: 32
#> Columns: 2
#> $ x <dbl> -3.8076427, -2.6742223, -3.8076427, -2.6742223, -3.8076427, -2.67422~
#> $ y <dbl> -6.2798254, -4.4744481, -2.6690708, -0.8636935, 0.9416838, 2.7470611~
\end{verbatim}

@@

\begin{verbatim}
hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = NA)
glimpse(hex_grid)
\end{verbatim}

\begin{verbatim}
#> Rows: 192
#> Columns: 3
#> $ x  <dbl> -2.674222, -2.674222, -3.807643, -4.941063, -4.941063, -3.807643, -~
#> $ y  <dbl> -5.6804828, -6.8791681, -7.4785108, -6.8791681, -5.6804828, -5.0811~
#> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
\end{verbatim}

\begin{verbatim}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red")
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-12-1.pdf}

@@

ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID))
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-13-1.pdf}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Assign data into hexagons
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Compute standardized counts
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{5}
\tightlist
\item
  Extract full grid info
\end{enumerate}

@@

\begin{verbatim}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue")
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-18-1.pdf}

\begin{verbatim}
ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-19-1.pdf}

@@

\begin{verbatim}
#>             x          y hexID std_counts
#> 1  -2.6742223 -4.4744481     5     1.0000
#> 2  -1.5408019 -6.2798254     2     0.3125
#> 3  -0.4073814 -4.4744481     6     0.0625
#> 4  -1.5408019 -2.6690708    10     0.2500
#> 5  -0.4073814 -0.8636935    14     0.5000
#> 6   0.7260390 -2.6690708    11     0.1250
#> 7   1.8594594 -0.8636935    15     0.1875
#> 8   0.7260390  0.9416838    19     0.6250
#> 9   1.8594594  2.7470611    23     0.2500
#> 10  0.7260390  4.5524384    27     0.5625
#> 11  1.8594594  6.3578158    31     0.3750
#> 12  2.9928798  4.5524384    28     0.4375
\end{verbatim}

@@

\begin{verbatim}
all_centroids_df_shift <- extract_coord_of_shifted_hex_grid(nldr_df = s_curve_noise_umap, 
                                                 x = "UMAP1", y = "UMAP2", 
                                                 num_bins_x = num_bins_x, 
                                                 num_bins_y = num_bins_y,
                                                 shift_x = 0.2690002, shift_y = 0.271183,
                                                 buffer_size = NA, hex_size = NA)

glimpse(all_centroids_df_shift)
\end{verbatim}

\begin{verbatim}
#> Rows: 32
#> Columns: 2
#> $ x <dbl> -3.8076427, -2.6742223, -3.8076427, -2.6742223, -3.8076427, -2.67422~
#> $ y <dbl> -6.2798254, -4.4744481, -2.6690708, -0.8636935, 0.9416838, 2.7470611~
\end{verbatim}

\begin{verbatim}
hex_grid <- gen_hex_coordinates(all_centroids_df_shift)
glimpse(hex_grid)
\end{verbatim}

\begin{verbatim}
#> Rows: 192
#> Columns: 3
#> $ x  <dbl> -2.674222, -2.674222, -3.807643, -4.941063, -4.941063, -3.807643, -~
#> $ y  <dbl> -5.6804828, -6.8791681, -7.4785108, -6.8791681, -5.6804828, -5.0811~
#> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
\end{verbatim}

\begin{verbatim}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = all_centroids_df_shift, aes(x = x, y = y), color = "red")
\end{verbatim}

\begin{verbatim}
full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df_shift)

ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID))
\end{verbatim}

\begin{verbatim}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue")
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-31-1.pdf}

\begin{verbatim}
ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-32-1.pdf}

\begin{verbatim}
#>             x          y hexID std_counts
#> 1  -2.6742223 -4.4744481     5     1.0000
#> 2  -1.5408019 -6.2798254     2     0.3125
#> 3  -0.4073814 -4.4744481     6     0.0625
#> 4  -1.5408019 -2.6690708    10     0.2500
#> 5  -0.4073814 -0.8636935    14     0.5000
#> 6   0.7260390 -2.6690708    11     0.1250
#> 7   1.8594594 -0.8636935    15     0.1875
#> 8   0.7260390  0.9416838    19     0.6250
#> 9   1.8594594  2.7470611    23     0.2500
#> 10  0.7260390  4.5524384    27     0.5625
#> 11  1.8594594  6.3578158    31     0.3750
#> 12  2.9928798  4.5524384    28     0.4375
\end{verbatim}

\begin{verbatim}
bin_centroids_shift <- ggplot(data = hex_full_count_df, aes(x = c_x, y = c_y)) +
  geom_point(color = "#bdbdbd") +
  geom_point(data = shifted_hex_coord_df, aes(x = c_x, y = c_y), color = "#feb24c") +
  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) +
  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

hex_grid_shift <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
  geom_polygon(fill = NA, color = "#feb24c", aes(group = polygon_id)) +
  geom_polygon(data = hex_full_count_df, aes(x = x, y = y, group = polygon_id),
               fill = NA, color = "#bdbdbd") +
  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) +
  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

## Before shift
before_shift_plot <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  coord_equal() +
  theme_void() +
  theme(legend.position="bottom", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) +
  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

## After shift
after_shift_plot <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  coord_equal() +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) +
  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
\end{verbatim}

@@

\begin{verbatim}
## As an option first quantile considered as a default
benchmark_to_rm_lwd_hex <- quantile(df_bin_centroids$std_counts)[2] + 0.01

## To identify low density hexagons
df_bin_centroids_low <- df_bin_centroids |>
  dplyr::filter(std_counts <= benchmark_to_rm_lwd_hex)

## To identify low-density hexagons needed to remove by investigating neighbouring mean density
identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_all = df_bin_centroids, num_bins_x = num_bins_x,
                     df_bin_centroids_low = df_bin_centroids_low)
\end{verbatim}

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-37-1.pdf}

@@

\begin{verbatim}
#> [1] 2.603
\end{verbatim}

\begin{verbatim}
benchmark <- 5
\end{verbatim}

@@

\begin{verbatim}
fit_high_d_model(training_data = training_data, nldr_df_with_id = s_curve_noise_umap, x = "UMAP1",
                             y = "UMAP1", num_bins_x = NA, num_bins_y = NA,
                             hex_size = NA, buffer_size = NA,
                             is_bin_centroid = TRUE,
                             is_rm_lwd_hex = FALSE,
                             benchmark_to_rm_lwd_hex = NA,
                             is_avg_high_d = TRUE, column_start_text = "x")
\end{verbatim}

\begin{verbatim}
#> $df_bin
#> # A tibble: 6 x 8
#>   hb_id     x1    x2     x3        x4        x5        x6        x7
#>   <int>  <dbl> <dbl>  <dbl>     <dbl>     <dbl>     <dbl>     <dbl>
#> 1     1 -0.470 0.473 -1.75   0.00121   0.00244  -0.0314   -0.000525
#> 2     2  0.170 1.23  -1.76   0.00610  -0.00313   0.00286  -0.00123 
#> 3     6  0.760 1.19  -0.407 -0.000154  0.00342   0.0219    0.000410
#> 4    11 -0.452 1.17   0.294  0.00414  -0.000295  0.00334   0.00207 
#> 5    14 -0.613 1.69   1.79   0.0119   -0.00418  -0.000519 -0.00251 
#> 6    15  0.366 0.999  1.70  -0.00115   0.00375  -0.00747  -0.000931
#> 
#> $df_bin_centroids
#>            x         y hexID std_counts
#> 1 -3.8076427 -3.807643     1 0.41666667
#> 2 -1.5408019 -3.807643     2 0.62500000
#> 3 -0.4073814 -1.540802     6 0.41666667
#> 4 -0.4073814  2.992880    14 0.08333333
#> 5  0.7260390  0.726039    11 0.58333333
#> 6  1.8594594  2.992880    15 1.00000000
\end{verbatim}

@@

\begin{verbatim}
#> Rows: 75
#> Columns: 4
#> $ pred_UMAP_1 <dbl> -2.6742223, 1.8594594, 0.7260390, -1.5408019, -1.5408019, ~
#> $ pred_UMAP_2 <dbl> -4.4744481, -0.8636935, 0.9416838, -6.2798254, -2.6690708,~
#> $ ID          <int> 1, 2, 3, 4, 6, 7, 8, 9, 11, 12, 14, 15, 16, 17, 19, 20, 21~
#> $ pred_hb_id  <int> 5, 15, 19, 2, 10, 31, 15, 11, 15, 28, 11, 5, 28, 5, 11, 23~
\end{verbatim}

@@

\begin{verbatim}
#> $mse
#> [1] 0.2983091
#> 
#> $aic
#> [1] -467.0531
\end{verbatim}

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-44-1.pdf}

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-45-1.pdf}

@@

\begin{verbatim}
medlea_df <- read_csv("data/medlea_dataset.csv")
names(medlea_df)[2:(NCOL(medlea_df) - 1)] <- paste0("x", 1:(NCOL(medlea_df) - 2))

medlea_df <- medlea_df |> ## Since only contains zeros
  select(-x10)

#medlea_df[,2:(NCOL(medlea_df) - 1)] <- scale(medlea_df[,2:(NCOL(medlea_df) - 1)])

calculate_pca <- function(feature_dataset, num_pcs){
  pcaY_cal <- prcomp(feature_dataset, center = TRUE, scale = TRUE)
  PCAresults <- data.frame(pcaY_cal$x[, 1:num_pcs])
  summary_pca <- summary(pcaY_cal)
  var_explained_df <- data.frame(PC= paste0("PC",1:50),
                               var_explained=(pcaY_cal$sdev[1:50])^2/sum((pcaY_cal$sdev[1:50])^2))
  return(list(prcomp_out = pcaY_cal,pca_components = PCAresults, summary = summary_pca, var_explained_pca  = var_explained_df))
}
features <- medlea_df[,2:(NCOL(medlea_df) - 1)]
pca_ref_calc <- calculate_pca(features, 8) 
pca_ref_calc$summary
\end{verbatim}

\begin{verbatim}
#> Importance of components:
#>                           PC1    PC2    PC3     PC4     PC5     PC6     PC7
#> Standard deviation     3.1691 3.0609 2.7226 1.87967 1.71219 1.34192 1.27525
#> Proportion of Variance 0.1969 0.1837 0.1453 0.06928 0.05748 0.03531 0.03189
#> Cumulative Proportion  0.1969 0.3806 0.5260 0.59526 0.65274 0.68805 0.71993
#>                            PC8     PC9    PC10    PC11    PC12    PC13   PC14
#> Standard deviation     1.16992 1.13465 1.06628 1.03279 0.97899 0.96264 0.9528
#> Proportion of Variance 0.02684 0.02524 0.02229 0.02091 0.01879 0.01817 0.0178
#> Cumulative Proportion  0.74677 0.77202 0.79431 0.81522 0.83402 0.85219 0.8700
#>                          PC15   PC16    PC17    PC18    PC19    PC20    PC21
#> Standard deviation     0.9116 0.9090 0.79750 0.76725 0.72414 0.65310 0.61052
#> Proportion of Variance 0.0163 0.0162 0.01247 0.01154 0.01028 0.00836 0.00731
#> Cumulative Proportion  0.8863 0.9025 0.91496 0.92650 0.93678 0.94514 0.95245
#>                          PC22    PC23    PC24    PC25    PC26   PC27    PC28
#> Standard deviation     0.6019 0.55399 0.52293 0.46638 0.41959 0.3976 0.34697
#> Proportion of Variance 0.0071 0.00602 0.00536 0.00426 0.00345 0.0031 0.00236
#> Cumulative Proportion  0.9596 0.96557 0.97093 0.97520 0.97865 0.9818 0.98411
#>                           PC29    PC30    PC31    PC32    PC33    PC34    PC35
#> Standard deviation     0.33415 0.30618 0.29237 0.28458 0.26033 0.25420 0.22792
#> Proportion of Variance 0.00219 0.00184 0.00168 0.00159 0.00133 0.00127 0.00102
#> Cumulative Proportion  0.98630 0.98814 0.98982 0.99140 0.99273 0.99400 0.99502
#>                           PC36    PC37    PC38   PC39    PC40    PC41    PC42
#> Standard deviation     0.21644 0.20437 0.19127 0.1744 0.15586 0.15252 0.12519
#> Proportion of Variance 0.00092 0.00082 0.00072 0.0006 0.00048 0.00046 0.00031
#> Cumulative Proportion  0.99594 0.99676 0.99747 0.9981 0.99855 0.99900 0.99931
#>                           PC43    PC44    PC45    PC46    PC47    PC48    PC49
#> Standard deviation     0.10485 0.08598 0.08008 0.06491 0.04841 0.04094 0.03791
#> Proportion of Variance 0.00022 0.00014 0.00013 0.00008 0.00005 0.00003 0.00003
#> Cumulative Proportion  0.99952 0.99967 0.99980 0.99988 0.99992 0.99996 0.99999
#>                           PC50    PC51
#> Standard deviation     0.02347 0.01421
#> Proportion of Variance 0.00001 0.00000
#> Cumulative Proportion  1.00000 1.00000
\end{verbatim}

\begin{verbatim}
var_explained_df <- pca_ref_calc$var_explained_pca
data_pca <- pca_ref_calc$pca_components |>
  mutate(ID = 1:NROW(pca_ref_calc$pca_components),
         shape_label = medlea_df$Shape_label)

var_explained_df |>
  ggplot(aes(x = PC,y = var_explained, group = 1))+
  geom_point(size=1)+
  geom_line()+
  labs(title="Scree plot: PCA on scaled data") +
  scale_x_discrete(limits = paste0(rep("PC", 50), 1:50)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-46-1.pdf}

\begin{verbatim}
data_split <- initial_split(data_pca)
training_data <- training(data_split) |>
  arrange(ID)
test_data <- testing(data_split) |>
  arrange(ID)
\end{verbatim}

\begin{verbatim}
UMAP_fit <- umap(training_data |> dplyr::select(-c(ID, shape_label)), n_neighbors = 37, n_components =  2)

UMAP_data <- UMAP_fit$layout |>
  as.data.frame()
names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))

UMAP_data <- UMAP_data |>
  mutate(ID = training_data$id)

UMAP_data_with_label <- UMAP_data |>
  mutate(shape_label = training_data$shape_label)
\end{verbatim}

\begin{verbatim}
UMAP_data_with_label |>
    ggplot(aes(x = UMAP1,
               y = UMAP2, color = shape_label))+
    geom_point(alpha=0.5) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-48-1.pdf}

\begin{verbatim}
tSNE_data <- Fit_tSNE(training_data |> dplyr::select(-c(ID, shape_label)), opt_perplexity = calculate_effective_perplexity(training_data |> dplyr::select(-c(ID, shape_label))), with_seed = 20240110)

tSNE_data <- tSNE_data |>
  select(-ID) |>
  mutate(ID = training_data$ID)

tSNE_data_with_label <- tSNE_data |>
  mutate(shape_label = training_data$shape_label)

>>>>

\begin{verbatim}
library(scales)
s_curve_noise_umap$UMAP1 <- rescale(s_curve_noise_umap$UMAP1)

y_min <- -sqrt(3)/2
y_max <- sqrt(3)/2

s_curve_noise_umap$UMAP2 <- ((s_curve_noise_umap$UMAP2 - min(s_curve_noise_umap$UMAP2))/(max(s_curve_noise_umap$UMAP2) - min(s_curve_noise_umap$UMAP2))) * (y_max - y_min)

# Define the desired size of the hexagons
hex_size <- 0.2

# Compute the horizontal spacing (dx) and vertical spacing (dy) between hexagon centers
dx <- sqrt(3) * hex_size
dy <- 1.5 * hex_size

x_length <- 1
y_length_scaled <- diff(range(s_curve_noise_umap$UMAP2))

# Compute the number of bins along the x-axis and y-axis
# num_bins_x <- ceiling(x_length / dx)
# num_bins_y <- ceiling(y_length_scaled / dy)
\end{verbatim}

\begin{verbatim}
# s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)] <- scale(s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)])
\end{verbatim}

@@

@@

\begin{verbatim}
num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap,
                                         x = "UMAP1", hex_size = 0.2)
num_bins_x
\end{verbatim}

\begin{verbatim}
#> [1] 3
\end{verbatim}

\begin{verbatim}
num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap,
                                         y = "UMAP2", hex_size = 0.2)
num_bins_y
\end{verbatim}

\begin{verbatim}
#> [1] 6
\end{verbatim}

@@

\begin{itemize}
\item
  First compute hex grid bound values along the x and y axis and generate the all the points within the hex box
\item
  Second for each x-value, find which y values are in the even row
\item
  Then, shift the x values of the even rows
\end{itemize}

\begin{verbatim}
all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
                                                 x = "UMAP1", y = "UMAP2", 
                                                 num_bins_x = num_bins_x, 
                                                 num_bins_y = num_bins_y, 
                                                 buffer_x = NA, buffer_y = NA, 
                                                 hex_size = 0.2)

@@

\begin{verbatim}
#> Rows: 18
#> Columns: 2
#> $ x <dbl> 0.1464102, 0.4928203, 0.8392305, 0.3196152, 0.6660254, 1.0124356, 0.~
#> $ y <dbl> 0.1845299, 0.1845299, 0.1845299, 0.4845299, 0.4845299, 0.4845299, 0.~
\end{verbatim}

@@

\begin{verbatim}
hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = 0.2)
glimpse(hex_grid)
\end{verbatim}

\begin{verbatim}
#> Rows: 108
#> Columns: 3
#> $ x  <dbl> 0.31961524, 0.31961524, 0.14641016, -0.02679492, -0.02679492, 0.146~
#> $ y  <dbl> 0.28412287, 0.08493702, -0.01465590, 0.08493702, 0.28412287, 0.3837~
#> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
\end{verbatim}

\begin{verbatim}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red") +
  coord_fixed()
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-10-1.pdf}

@@

ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID)) +
  coord_fixed()
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Assign data into hexagons
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{5}
\tightlist
\item
  Compute standardized counts
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{6}
\tightlist
\item
  Extract full grid info
\end{enumerate}

@@

\begin{verbatim}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue") +
  coord_fixed()
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-16-1.pdf}

\begin{verbatim}
ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
  coord_fixed()
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-17-1.pdf}

@@

\begin{verbatim}
#>            x         y hexID std_counts
#> 1  0.1464102 0.1845299     1 1.00000000
#> 2  0.4928203 0.1845299     2 0.04761905
#> 3  0.3196152 0.4845299     4 0.19047619
#> 4  0.6660254 0.4845299     5 0.23809524
#> 5  0.4928203 0.7845299     8 0.19047619
#> 6  0.8392305 0.7845299     9 0.33333333
#> 7  0.6660254 1.0845299    11 0.28571429
#> 8  1.0124356 1.0845299    12 0.04761905
#> 9  0.8392305 1.3845299    15 0.38095238
#> 10 0.6660254 1.6845299    17 0.47619048
#> 11 1.0124356 1.6845299    18 0.38095238
\end{verbatim}

@@

@@

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-21-1.pdf}

@@

\begin{verbatim}
#> [1] 0.6
\end{verbatim}

@@

@@

\begin{verbatim}
#> Rows: 75
#> Columns: 4
#> $ pred_UMAP_1 <dbl> 0.1464102, 0.8392305, 1.0124356, 0.1464102, 0.3196152, 0.6~
#> $ pred_UMAP_2 <dbl> 0.1845299, 0.7845299, 1.0845299, 0.1845299, 0.4845299, 1.6~
#> $ ID          <int> 1, 2, 3, 4, 6, 7, 8, 9, 11, 12, 14, 15, 16, 17, 19, 20, 21~
#> $ pred_hb_id  <int> 1, 9, 12, 1, 4, 17, 9, 5, 9, 18, 5, 1, 18, 1, 5, 15, 2, 15~
\end{verbatim}

@@

\begin{verbatim}
#> $mse
#> [1] 0.372034
#> 
#> $aic
#> [1] -365.1042
\end{verbatim}

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-27-1.pdf}

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-28-1.pdf}

\hypertarget{find-non-empty-bins}{%
\subsubsection{Find non-empty bins}\label{find-non-empty-bins}}

@@
